---
title: "CaseStudy#1"
author: "Veysel Bugra Demir"
date: "01 06 2022"
output: 
  html_document: 
    toc: yes
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Google Divvy Case Study

Required packeges
```{r}
library(tidyverse) 
library(lubridate)  
library(ggplot2)  
```

# STEP 1: Data Collection
```{r}
q1_2019 <- read_csv("C:\\Users\\Buğra\\Desktop\\bike data\\data\\Divvy_Trips_2019_Q1.csv")
q2_2019 <- read_csv("C:\\Users\\Buğra\\Desktop\\bike data\\data\\Divvy_Trips_2019_Q2.csv")
q3_2019 <- read_csv("C:\\Users\\Buğra\\Desktop\\bike data\\data\\Divvy_Trips_2019_Q3.csv")
q4_2019 <- read_csv("C:\\Users\\Buğra\\Desktop\\bike data\\data\\Divvy_Trips_2019_Q4.csv")

```
# STEP 2: Wrangle and Combine
Renaming the columns for consistency.
```{r}
q1_2019 <- rename(q1_2019, ride_id=trip_id,
                  rideable_type=bikeid,
                  started_at = start_time,
                  ended_at = end_time,
                  start_station_name = from_station_name,
                  start_station_id = from_station_id,
                  end_station_name = to_station_name,
                  end_station_id = to_station_id,
                  member_casual = usertype)

q2_2019 <- rename(q2_2019, ride_id="01 - Rental Details Rental ID",
                  rideable_type="01 - Rental Details Bike ID",
                  started_at = "01 - Rental Details Local Start Time" ,
                  ended_at = "01 - Rental Details Local End Time",
                  start_station_name = "03 - Rental Start Station Name",
                  start_station_id = "03 - Rental Start Station ID",
                  end_station_name = "02 - Rental End Station Name",
                  end_station_id = "02 - Rental End Station ID",
                  member_casual = "User Type")

q3_2019 <- rename(q3_2019, ride_id=trip_id,
                  rideable_type=bikeid,
                  started_at = start_time,
                  ended_at = end_time,
                  start_station_name = from_station_name,
                  start_station_id = from_station_id,
                  end_station_name = to_station_name,
                  end_station_id = to_station_id,
                  member_casual = usertype)

q4_2019 <- rename(q4_2019, ride_id=trip_id,
                  rideable_type=bikeid,
                  started_at = start_time,
                  ended_at = end_time,
                  start_station_name = from_station_name,
                  start_station_id = from_station_id,
                  end_station_name = to_station_name,
                  end_station_id = to_station_id,
                  member_casual = usertype)
```

```{r}
str(q1_2019)
str(q2_2019)
str(q3_2019)
str(q4_2019)

```
Converting data.
```{r}
q1_2019 <- mutate(q1_2019, ride_id = as.character(ride_id), rideable_type = as.character(rideable_type))

q2_2019 <- mutate(q2_2019, ride_id = as.character(ride_id), rideable_type = as.character(rideable_type))

q3_2019 <- mutate(q3_2019, ride_id = as.character(ride_id), rideable_type = as.character(rideable_type))

q4_2019 <- mutate(q4_2019, ride_id = as.character(ride_id), rideable_type = as.character(rideable_type))
```
Concatenating the all data.
```{r}
all_trips <- bind_rows(q1_2019, q2_2019, q3_2019, q4_2019)
```
Dropping unnecessary columns.
```{r}
all_trips <- all_trips %>% 
  select(-c(birthyear, gender,"01 - Rental Details Duration In Seconds Uncapped", "05 - Member Details Member Birthday Year", "Member Gender", "tripduration"))
```
# STEP 3: CLEAN UP
```{r}
colnames(all_trips)
nrow(all_trips)
dim(all_trips)
head(all_trips) 
str(all_trips)
summary(all_trips)
```
Replacing values in the "member_casual" column
```{r}
table(all_trips$member_casual) # How many observations.

```
```{r}
all_trips <-all_trips %>% mutate(member_casual = recode(member_casual, "Subscriber" = "member",
                                            "Customer" = "casual"))
```
```{r}
table(all_trips$member_casual)
```
Creating new date, month, day, year and day of week columns to aggregate the results more granular level.
```{r}
all_trips$date <- as.Date(all_trips$started_at)
all_trips$month <- format(as.Date(all_trips$date), "%m")
all_trips$day <- format(as.Date(all_trips$date), "%d")
all_trips$year <- format(as.Date(all_trips$date), "%Y")
all_trips$day_of_week <- format(as.Date(all_trips$date), "%A")
```
Calculating trip durations.
```{r}
all_trips$ride_length <- difftime(all_trips$ended_at, all_trips$started_at)
```
```{r}
str(all_trips)
```
day_of_week was returned in Turkish
to make it in English use:
Sys.setlocale("LC_TIME","English United States")
```{r}
typeof(all_trips$ride_length)
```
Changing the data type for further calculations.
```{r}
all_trips$ride_length <- as.numeric(as.character(all_trips$ride_length))
```
typeof(all_trips$ride_length)
```{r}
typeof(all_trips$ride_length)
is.numeric(all_trips$ride_length)
```
Deleting the rows that has minus ride length.  
```{r}
all_trips_v2 <- all_trips[!(all_trips$start_station_name == "HQ QR" | all_trips$ride_length<0),]
```
# STEP 4: Analysis
```{r}
mean(all_trips_v2$ride_length)
median(all_trips_v2$ride_length)
max(all_trips_v2$ride_length)
min(all_trips_v2$ride_length)
```
```{r}
summary(all_trips_v2$ride_length)
```
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = mean) # is as the same as all_trips_v2 %>% group_by(member_casual) %>% summarise(mean(ride_length))
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = median)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = max)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = min)
```
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN=mean)
```
```{r}
all_trips_v2$day_of_week <- ordered(all_trips_v2$day_of_week, levels=c("Pazar", "Pazartesi", "Salı", "Çarşamba", "Perşembe", "Cuma", "Cumartesi"))
```
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)
```
```{r}
all_trips_v2 %>% 
  mutate(weekday= wday(started_at, label=TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides = n(), average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)
```
Visualize the data
```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides = n(), average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday) %>% 
  ggplot(aes(x=weekday, y=average_duration)) + geom_col(position="dodge", aes(fill=member_casual))
```
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)

```